---
title: "ML_Group_Project"
output: pdf_document
date: "2024-11-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
library(boot)
library(MASS)
```

```{r, echo=FALSE}
test_data = read.csv("test.csv")
train_data = read.csv("train.csv")
```

# Data Exploration

ADD SAMARA'S part

```{r}
#cor(df_train[,-c(1,10)])
```


It is worth to note that "decane_toluene" is not in the test data set


# Missing Data

Missing data was found in the 'parentspecies' attribute. According to the definition of the 'parentspecies' attribute, missing values imply a meaning.They are not missing randomly but due to difficulty in retrieving the 'parentspecies'. Therefore, a new level was created as 'Unknown'. 

```{r}

test_data$parentspecies[test_data$parentspecies == ""] <- "Unknown"
train_data$parentspecies[train_data$parentspecies == ""] <- "Unknown"

```


```{r, echo=FALSE}
train_data$parentspecies = factor(train_data$parentspecies)
train_levels <- levels(train_data$parentspecies)  # Get levels from train data
# Apply those levels to the test data
test_data$parentspecies <- factor(test_data$parentspecies, levels = train_levels)
```


# Dummy model


A dummy model is a supervised learning model that gives the same constant output regardless of the values of the covariates.
We first built a dummy model and calculated the training error and the cross validation error.



```{r}
set.seed(123)
dummy_model = glm(log_pSat_Pa ~ 1, data = train_data[,-1])
dummy_cv_error_5 = cv.glm(train_data[,-1] , dummy_model , K = 10)$delta[1]
dummy_error_train = mean((train_data$log_pSat_Pa - predict(dummy_model, train_data))^2)
```

```{r, echo=FALSE}
error_df = data.frame(model = "Dummy", train = dummy_error_train, cv = dummy_cv_error_5, kaggle_score = -0.0001)
knitr::kable(error_df)
```

# OLS as a baseline model

ten fold cross validation was done to get the cross validation error

```{r, warning=FALSE}
set.seed(123)
ols_fit = glm(log_pSat_Pa ~ ., data = train_data[,-1])
cv_error_10 = cv.glm(train_data[,-1] , ols_fit , K = 10)$delta[1]
error_train = mean((train_data$log_pSat_Pa - predict(ols_fit, train_data))^2)

```

# Plotting

```{r, echo=FALSE}
#summary(ols_fit)
par(mfrow=c(2,2))
plot(ols_fit)

hist(ols_fit$residuals)
```


```{r, echo=FALSE}
new_row <- data.frame(model = "ols", train = error_train, cv = cv_error_5, kaggle_score = 0.7163)
error_df <- rbind(error_df, new_row)
knitr::kable(error_df)
```



```{r, echo=FALSE}
y_hat = predict(ols_fit,newdata = test_data)

#ID = test_data$ID
#TARGET = as.vector(y_hat)
#df = data.frame(ID, TARGET)
#write.csv(df[, c("ID", "TARGET")], "dummy_submission.csv", row.names = FALSE)
```


## Then we analyze the correlation between each column and the target, selecting columns with a correlation above 0.5 or 0.3, and then using these for the model prediction. However, even after trying these two methods, the prediction score didn’t change,

# Lasso 

Before moving to non-linear models we tried regularization using Lasso

"decane_toluene" is not in the test data set. so it was handled in a way so that both train and test dataset has the same number of levels

```{r, echo=FALSE}
a1 = table(train_data$parentspecies)
a2 = table(test_data$parentspecies)

df = data.frame(a1,a2)
df2 = df[,-3]
colnames(df2) <- c("Level", "Freq_train", "Freq_test")

knitr::kable(df2)
```


```{r, echo=FALSE}
library(glmnet)

y <- train_data$log_pSat_Pa  

x_train <- model.matrix(log_pSat_Pa ~ . - 1, data = train_data[,-1])

lasso_model <- glmnet(x_train, y, family = "gaussian", alpha = 1)

error_train_lasso = mean((train_data$log_pSat_Pa - predict(lasso_model, x_train))^2)

```

Cross validation to obtain the best lambda(tuning parameter)

```{r, echo=FALSE}
set.seed(123)
cvfit = cv.glmnet(x_train, y, family = "gaussian", type.measure = "mse")


best_lambda = cvfit$lambda.min
#best_lambda

```

best lambda obtained was 0.002651092

Plot
```{r}
plot(cvfit)
```

WE can use lasso as a variable selection method

```{r, echo=FALSE}
lasso_coefficients = coef(cvfit, s = "lambda.min")
non_zero_coeffs = lasso_coefficients[lasso_coefficients != 0]
c = predict(lasso_model,type = "coefficients", s = best_lambda)[1:26,]
knitr::kable(c)
```



```{r, echo=FALSE}
error_cv_lasso = mean((train_data$log_pSat_Pa - predict(cvfit, x_train, s = best_lambda))^2)

new_row2 <- data.frame(model = "Lasso", train = error_train_lasso, cv = error_cv_lasso, kaggle_score = 0.7160)
error_df <- rbind(error_df, new_row2)

knitr::kable(error_df)

```

```{r, echo=FALSE}
x_test <- model.matrix(~ . - 1, data = test_data[,-1])
y_pred_test = predict(cvfit, newx = x_test, s = best_lambda)
```



```{r, echo=FALSE}
#ID = test_data$ID
#TARGET = as.vector(y_pred_test)
#df = data.frame(ID, TARGET)
#write.csv(df[, c("ID", "TARGET")], "dummy_submission.csv", row.names = FALSE)
```

PCR

# Regression tree

```{r}

```


Boosting 

And then gradient Boosting by coco

Random forests

SVR


